---
title: "Analyses"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r load-libraries, message = FALSE}
x <- c("vroom", "MKinfer", "ff", "car", "lme4", "boot", "R.utils", "dplyr", "tidyverse", "data.table", "lme4", "leaps", "glmnet")
invisible(lapply(x, library, character.only = TRUE))
```

```{r load-data}
data_dir <- 'C:/Users/Roderick/Box Sync/Research Projects/Variability Adaptation/open-data/adaptation_skill_acquisition/data'
# Set data_dir to location of data files
setwd(data_dir)

human_per_game <- fread('human_per_game.csv')
follow_up_per_game <- fread('follow_up_per_game.csv')
models_slowdown_per_game <- fread('models_slowdown_per_game.csv')
models_noslow_per_game <- fread('models_noslow_per_game.csv')
models_decay_per_game <- fread('models_decay_per_game.csv')

```


Comparing the effect of varied vs. consistent training on points in block 5 using multiple regression
```{r varied-consistent-comparison}

human_per_game %>%
  filter(experiment==1, block==5, condition!="MMMMM") %>%
  group_by(participant, varied, switch.type) %>%
  summarise(raw.points = mean(raw.points)) %>%
  lm(raw.points ~ as.factor(varied)*as.factor(switch.type), data=.) %>%
  summary()

human_per_game %>%
  filter(experiment==2, block==5, condition!="HHHHH") %>%
  group_by(participant, varied, switch.type) %>%
  summarise(raw.points = mean(raw.points)) %>%
  lm(raw.points ~ as.factor(varied)*as.factor(switch.type), data=.) %>%
  summary() 

```
Comparing the effect of blocked vs. interleaved training on points in block 5 using t-test
```{r follow-up-comparison}
follow_transfer <- follow_up_per_game %>%
  group_by(participant, condition, experiment) %>%
  filter(trial_num > 32) %>%
  summarise(raw_points = mean(raw_points)) %>%
  mutate(interleaved = case_when(condition %in% c("LH16M", "HL16M") ~ 1,
                                 TRUE ~ 0)) %>%
  ungroup()


t.test(follow_transfer[follow_transfer$interleaved==0, 'raw_points'],
       follow_transfer[follow_transfer$interleaved==1, 'raw_points'])

sd(as.numeric(unlist(follow_transfer[follow_transfer$interleaved==1, 'raw_points'])))

```

Estimating practice and acceleration effects
```{r regression-practice-acceleration}

regression_practice_acceleration <- function(organism){

  data <- NULL
  if(organism =='human'){
    data <- human_per_game
  }
  else{
    data <- models_decay_per_game %>%
      filter(decay.value==0.992)
  }
  
  lmm.model <- data %>%
    mutate(trial.num.in.block = ((trial.num-1) %% 8) + 1) %>% 
    group_by(participant, log.block, acceleration) %>%
    summarize(raw.points = mean(raw.points)) %>%
    lmer(raw.points ~ log.block + acceleration + (1 | participant), data = ., REML = FALSE)


  return(lmm.model)
}


regression_practice_acceleration('human')
regression_practice_acceleration('model')

```

Estimating switch-faster and switch-slower effects
```{r}

regression_switch <- function(organism){
  
  data <- NULL
  if(organism =='human'){
    data <- human_per_game
  }
  else{
    data <- models_decay_per_game %>%
      filter(decay.value==0.992)
  }
  
  lmm.model <- regression_practice_acceleration(organism)
  
  # Extract coefficients for block,  acceleration, and individual effects
  # Adjust points using extracted coefficients
  lmm.model.coef <- coef(lmm.model)$participant
  names(lmm.model.coef) <- c("sub.coef", "log.block.coef", "acceleration.coef")
  lmm.model.coef <- tibble::rownames_to_column(lmm.model.coef, "participant")
  updated.data.all <- left_join(x=data,y=lmm.model.coef) %>%
    mutate(adjust.raw.points = raw.points - sub.coef - log.block.coef*log.block - 
             acceleration.coef*acceleration)
  
  
  lm.adjusted.switch <- updated.data.all %>%
    mutate(trial.num.in.block = ((trial.num-1) %% 8) + 1,
           switch.slower = -1*switch.slower) %>% 
    filter(trial.num.in.block==1, block!=1) %>%
    group_by(participant, switch.faster, switch.slower, block) %>%
    summarize(adjust.raw.points = mean(adjust.raw.points)) %>%
    lm(adjust.raw.points ~ switch.faster + switch.slower, data = .)
  
  return(lm.adjusted.switch)
}

summary(regression_switch('human'))
summary(regression_switch('model'))

```

Comparing switch-faster and switch-slower effects using a F test
```{r switch-effect-comparison}

linearHypothesis(regression_switch('human'), "switch.slower = switch.faster", test="F", verbose=TRUE)
linearHypothesis(regression_switch('model'), "switch.slower = switch.faster", test="F", verbose=TRUE)
```

Model fits to human data
```{r model-fits}

# Merge human with model data
first_fit_data <- human_per_game %>%
    mutate(bad.weight = 'human',
           decay.value = 'human',
           slowdown = 'human',
           minThrustTime = 'human',
           player.type = 'human') %>%
  rbind(., models_noslow_per_game, models_slowdown_per_game)

second_fit_data <- human_per_game %>%
    mutate(bad.weight = 'human',
           decay.value = 'human',
           slowdown = 'human',
           minThrustTime = 'human',
           player.type = 'human') %>%
  rbind(., models_decay_per_game)

# Chi square function
chi_fits_avg_se <- function(data, var){
    myvar <- enquo(var)

    temp <- data %>%
        group_by(trial.num, condition, player.type) %>%
        summarise(mean = mean(!!myvar))
        
    sq.se <- data %>% 
        filter(player.type == 'human') %>%
        group_by(trial.num, condition, player.type) %>%
        summarise(mean = mean(!!myvar), 
                  sq.std.error = (sd(!!myvar)**2)/length(!!myvar))

    avg.sq.se <- mean(sq.se$sq.std.error)
    
    temp_wide <- spread(temp, player.type, mean)

    temp_wide$trial.num <- NULL
    temp_wide$condition <- NULL
    
    myfun <- function(i){
        sum((i - temp_wide$human)**2)
    }
    
    SSE <- apply(temp_wide, 2, myfun)
    chi.sq <- SSE/avg.sq.se
    
    full_stats <- data.frame(chi.sq)
    colnames(full_stats) <- c(sprintf('chi.sq.%s', as_label(myvar)))
    setDT(full_stats, keep.rownames = 'player.type')[]
    return(full_stats)
}

# Fits for first set of model comparisons
first_fit_chis <- merge(chi_fits_avg_se(first_fit_data, rectangles.cleared), 
                      chi_fits_avg_se(first_fit_data, deaths)) %>%
  mutate(chi.sq.combined = chi.sq.rectangles.cleared + chi.sq.deaths) %>%
  na.omit(.)

# Fits for second set of model comparisons
second_fit_chis <- merge(chi_fits_avg_se(second_fit_data, rectangles.cleared), 
                      chi_fits_avg_se(second_fit_data, deaths)) %>%
  mutate(chi.sq.combined = chi.sq.rectangles.cleared + chi.sq.deaths) %>%
  na.omit(.)

first_fit_chis
second_fit_chis

```

